---
title: "BU662 - Project 5 - Trading with Quantmod
Name: Rishant Rokaha
---




Project:

1.	Pick 3 common stocks (see list), ETF's, currencies or other financial instruments. Check to see what business they are engaged in (i.e. products, services, markets). Compare the side by side (same graph) performance of the stocks since 1/1/2018. 
2.	In addition, compare your stocks with the fluctuations and returns you could have achieved by investing in the S&P 500 (GSPC) or gold (GLD).
3.	Using Quantmod charts and 2-3 of the common stock indicators (see list), determine which would have been most helpful in helping a trader make a profit by using any of the selected indicators.

Based upon your indicators when the trades (puts (sell) or calls (buy)) that should have been made to maximize profits.

Include the coding in "code chunks" or in a separate R script. Explain your trading strategy and time of trade execution. Explain why you did what you did in using the code.



List of Stocks

Google (GOOG)
Microsoft (MSFT)
International Business Machines (IBM)
AT&T (T)
Amazon (AMZN)
Goldman Sachs (GS)
Boeing (BA)
Bank of America (BAC)
Exxon-Mobil (XOM)
Lockheed – Martin (LMT)
Starbucks (SBUX)
Macy’s (M)
Wall-Mart  (WMT)
Federal Express (FDX)



List of Common Indicators:

addMACD()# adds moving average convergence divergence signals to the apple stock price

addBBands() # Adds Bollinger bands to the apple stock price.
## A Bollinger Band®, developed by famous technical trader John Bollinger, 
## is plotted two standard deviations away from a simple moving average.
## the closer the prices move to the upper band, the more overbought the market, 
## and the closer the prices move to the lower band, the more oversold the market

addCCI() # Add Commodity channel index.
## CCI measures the current price level relative to an average price level over a given period 
## of time. CCI is relatively high when prices are far above their average. 
##CCI is relatively low when prices are far below their average. 
## In this manner, CCI can be used to identify overbought and oversold levels. 
addADX() #Add Directional Movement Indicator
## Reading directional signals is easy. When the +DMI is dominant and rising, price direction is up. 
## When the -DMI is dominant and rising, price direction is down. But the strength ## of price 
## must also be considered. DMI strength ranges from a low of 0 to a high of 100. 
## The higher the DMI value, the stronger the prices swing. 
## DMI values over 25 mean price is directionally strong. 
## DMI values under 25 mean price is directionally weak.


addCMF() #Add Chaiken Money Flow
## Chaikin believed that buying and selling pressures could be determined by where a period closes 
## in relation to its high/low range. If the period closes in the upper half of the range, 
## then buying pressure is higher and if the period closes in the lower half of the range, 
## then selling pressure is higher. 

```{r}
## install.packages("quantmod")
library(quantmod)
library(Quandl)
require(TTR)
```


#Retrieve all the symbols required for the Company stocks and data from 2018 to 2020 

```{r}
getSymbols("^GSPC", src = "yahoo", from =  as.Date("2018-01-01"), to = as.Date("2020-04-30"))
tail(GSPC,10)
getSymbols("GLD", src = "yahoo", from =  as.Date("2018-01-01"), to = as.Date("2020-04-30"))
tail(GLD,10)
getSymbols("AMZN", src = "yahoo", from =  as.Date("2018-01-01"), to = as.Date("2020-04-30"))
tail(AMZN,10)
getSymbols("MSFT", src = "yahoo", from =  as.Date("2017-01-01"), to = as.Date("2020-04-30"))
tail(MSFT,10)
getSymbols("IBM", src = "yahoo", from =  as.Date("2017-01-01"), to = as.Date("2020-04-30"))
tail(IBM,10)


```


# We will create charts for all companies stocks for 2018.
```{r}
chartSeries(AMZN,subset="2018",theme="white")
chartSeries(MSFT,subset="2018",theme="white")
chartSeries(IBM,subset="2018",theme="white")
chartSeries(GLD,subset="2018",theme="white")
chartSeries(GSPC,subset="2018",theme="white")
```


#Here We designed our own theme so that our graph will be more visible
```{r}
whiteTheme <- chartTheme("white") 

## List of desig options in "theme" function
names(whiteTheme)

whiteTheme$bg.col <- "white"
whiteTheme$dn.col <- "red"
whiteTheme$up.col <- "green"
whiteTheme$border <- "lightgray"
a <- chartSeries(AMZN,subset="last 3 months",theme=whiteTheme,TA=NULL)
b <- chartSeries(MSFT,subset="last 3 months",theme=whiteTheme,TA=NULL)
c <- chartSeries(IBM,subset="last 3 months",theme=whiteTheme,TA=NULL)
d <- chartSeries(GLD,subset="last 3 months",theme=whiteTheme,TA=NULL)
e <- chartSeries(GSPC,subset="last 3 months",theme=whiteTheme,TA=NULL)
#class(a)
#class(b)
#class(c)
#class(d)
#class(e)
```

#Chart series using name and our own designed theme
```{r}
chartSeries(AMZN["2018"],theme=whiteTheme,name="Amazon")
chartSeries(MSFT["2018"],theme=whiteTheme,name="Microsoft")
chartSeries(IBM["2018"],theme=whiteTheme,name="IBM")
chartSeries(GLD["2018"],theme=whiteTheme,name="Gold")
chartSeries(GSPC["2018"],theme=whiteTheme,name="S&P 500")

```



#Print out all the monthly returns of 2018 for all companies to compare returns 
```{r}

returnAMZN<-monthlyReturn(AMZN, subset="2018") 
returnAMZN
returnMSFT<-monthlyReturn(MSFT, subset="2018") 
returnMSFT
returnIBM<-monthlyReturn(IBM, subset="2018") 
returnIBM
returnGLD<-monthlyReturn(GLD, subset="2018") 
returnGLD
returnGSPC<-monthlyReturn(GSPC, subset="2018") 
returnGSPC
```

#Create Chart Series for all company stocks

```{r}
chart_Series(Cl(GSPC), subset = "2018")
add_TA(Cl(AMZN), on = 1, col = "blue")
add_TA(Cl(MSFT), on = NA, col = "red")
add_TA(Cl(IBM), on = 1, col = "green")
add_TA(Cl(GLD), on = NA, col = "yellow")
```


# We normalized the data and put all graphs in the same chart to compare side by side
```{r}
normalise_series <- function(xdat) xdat / coredata(xdat)[1]
window <- "2018"

# Define colour of default chart line to chart_Series in mytheme object
# which is passed to chart_Series:
mytheme <- chart_theme()
mytheme$col$line.col <- "purple"
chart_Series(normalise_series(Cl(GSPC)[window]) - 1, theme = mytheme)
add_TA(normalise_series(Cl(AMZN)[window]) - 1, on = 1, col = "blue", lty = 3)
add_TA(normalise_series(Cl(MSFT)[window]) - 1, on = 1, col = "red", lty = 3)
add_TA(normalise_series(Cl(IBM)[window]) - 1, on = 1, col = "green", lty =2)
add_TA(normalise_series(Cl(GLD)[window]) - 1, on = 1, col = "black", lty = 3)
# Or add RSIs on different subpanels to improve readability of charts:
#add_TA(RSI(Cl(GSPC)), on = 1, col = "purple", lty = 2)
#add_TA(RSI(Cl(AMZN)), on =1, col = "blue")
#add_TA(RSI(Cl(MSFT)), on = 1, col = "red", lty = 3)

#add_TA(RSI(Cl(IBM)), on = 1, col = "green", lty = 2)
#add_TA(RSI(Cl(GLD)), on = 1, col = "black", lty = 2)
```

#Simple Moving averages comparison (SMA) for 20, 50 and 100 trading days.

```{r}
candleChart(AMZN,TA="addSMA(n=c(20,50,100))", up.col = "black", dn.col = "red", theme = whiteTheme, subset = "2018")


candleChart(MSFT,TA="addSMA(n=c(20,50,100))", up.col = "black", dn.col = "red", theme = whiteTheme, subset = "2018")


candleChart(IBM,TA="addSMA(n=c(20,50,100))", up.col = "black", dn.col = "red", theme = whiteTheme, subset = "2018")

```

#Moving Average Convergence-Divergence (MACD)
#addMACD() adds moving average convergence divergence signals to the stock price

```{r}
macd <- MACD( Cl(AMZN), 20, 26, 9, maType="EMA" ) 
tail(macd)
chartSeries(AMZN,TA = "addMACD()", subset="2018",theme =whiteTheme)


macd <- MACD( Cl(MSFT), 12, 26, 9, maType="EMA" ) 
tail(macd)
chartSeries(MSFT,TA = "addMACD()", subset="2018",theme=whiteTheme)


macd <- MACD( Cl(IBM), 12, 26, 9, maType="EMA" ) 
tail(macd)
chartSeries(IBM,TA = "addMACD()", subset="2018",theme=whiteTheme)

```


Bollinger Bands
#addBBands()  Adds Bollinger bands to the apple stock price.
## A Bollinger Band®, developed by famous technical trader John Bollinger, 
## is plotted two standard deviations away from a simple moving average.
## the closer the prices move to the upper band, the more overbought the market, 
## and the closer the prices move to the lower band, the more oversold the market
```{r}
bAMZN <- BBands(HLC=HLC(AMZN["2018"]), n=20, sd=2)
tail(bAMZN,10)
chartSeries(AMZN,TA='addBBands();addBBands(draw="p");addVo()', subset = "2018", theme=whiteTheme)


bMSFT <- BBands(HLC=HLC(MSFT["2018"]), n=20, sd=2)
tail(bMSFT,10)
chartSeries(MSFT,TA='addBBands();addBBands(draw="p");addVo()', subset = "2018", theme=whiteTheme)



bIBM <- BBands(HLC=HLC(IBM["2018"]), n=20, sd=2)
tail(bIBM,10)
chartSeries(IBM,TA='addBBands();addBBands(draw="p");addVo()', subset = "2018", theme = whiteTheme)

``` 


Commodity channel index (CCI): 
## addCCI()  Add Commodity channel index.
## CCI measures the current price level relative to an average price level over a given period 
## of time. CCI is relatively high when prices are far above their average. 
## CCI is relatively low when prices are far below their average. 
## In this manner, CCI can be used to identify overbought and oversold levels.
```{r}
chartSeries(AMZN, subset="2018",theme=whiteTheme)
addCCI(n = 20, maType="SMA", c=0.015)
chartSeries(MSFT, subset="2018",theme=whiteTheme)
addCCI(n = 20, maType="SMA", c=0.015)
chartSeries(IBM, subset="2018",theme=whiteTheme)
addCCI(n = 20, maType="SMA", c=0.015)
```

Directional Movement Indicator (ADX)
## addADX() Add Directional Movement Indicator
## Reading directional signals is easy. When the +DMI is dominant and rising, price direction is up. 
## When the -DMI is dominant and rising, price direction is down. But the strength ## of price 
## must also be considered. DMI strength ranges from a low of 0 to a high of 100. 
## The higher the DMI value, the stronger the prices swing. 
## DMI values over 25 mean price is directionally strong. 
## DMI values under 25 mean price is directionally weak.
```{r}
chartSeries(AMZN, subset="2018",theme=whiteTheme)
addADX(n = 14, maType="EMA", wilder=TRUE)
chartSeries(MSFT, subset="2018",theme=whiteTheme)
addADX(n = 14, maType="EMA", wilder=TRUE)
chartSeries(IBM, subset="2018",theme=whiteTheme)
addADX(n = 14, maType="EMA", wilder=TRUE)
```

Chaikin Money Flow (CMF)
## addCMF() Add Chaiken Money Flow
## Chaikin believed that buying and selling pressures could be determined by where a period closes 
## in relation to its high/low range. If the period closes in the upper half of the range, 
## then buying pressure is higher and if the period closes in the lower half of the range, 
## then selling pressure is higher. 

```{r}
candleChart(AMZN, up.col = "green", dn.col = "red", theme =whiteTheme, subset = "2018-01/")
addCMF()
candleChart(MSFT, up.col = "green", dn.col = "red", theme = whiteTheme, subset = "2018-01/")
addCMF()
candleChart(IBM, up.col = "green", dn.col = "red", theme = whiteTheme, subset = "2018-01/")
addCMF()
```
#Get Price Earnings ratio:
#Compared with Google
```{r}
getQuote("AMZN", what = yahooQF("P/E Ratio"))
getQuote("MSFT", what = yahooQF("P/E Ratio"))
getQuote("IBM", what = yahooQF("P/E Ratio"))
getQuote("GOOG", what = yahooQF("P/E Ratio"))
```

#Get Price Earnings Ratio
#Compared with Google, Apple, Netflix

```{r}
require(quantmod)
require("plyr")
what_metrics <- yahooQF(c("Price/Sales", 
                          "P/E Ratio",
                          "Price/EPS Estimate Next Year",
                          "PEG Ratio",
                          "Dividend Yield", 
                          "Market Capitalization"))

tickers <- c("AMZN", "MSFT", "IBM" , "GOOG","AAPL","NFLX")
# Not all the metrics are returned by Yahoo.
metrics <- getQuote(paste(tickers, sep="", collapse=";"), what=what_metrics)

#Add tickers as the first column and remove the first column which had date stamps
metrics <- data.frame(Symbol=tickers, metrics[,2:length(metrics)]) 

#Change colnames
colnames(metrics) <- c("Symbol", "P/E Ratio", "P/E EPS next year", "Div Yield", "Market Cap")

head(metrics,10)

```





